---
title: "0 - CORE - Data Import - R"
author: "Alister Palmer, Yujun He, Taleni Patterson, Dan Bridgman"
date: "28 September 2018"
output: html_document
---

# Setup & installing packages that maybe needed - To add new packages just add to list.of.packages vector
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(pastecs)
require(kableExtra)
require(corrplot)
require(ggplot2)
require(VIM)
require(caret)

```

# Reading in the data
```{r}

# Reading in the functions to import data.
source("C:\\Users\\dan bridgman\\Desktop\\Uni\\Projects\\STAT5003\\Code\\0 - PrepFunctions.R")

## TRAINING DATA ###
file_data_cli <- "C:\\Users\\dan bridgman\\Desktop\\Uni\\Projects\\STAT5003\\Data\\Group project\\train_cli.tsv"
file_data_pro <- "C:\\Users\\dan bridgman\\Desktop\\Uni\\Projects\\STAT5003\\Data\\Group project\\train_pro.tsv"
training_labels <- "C:\\Users\\dan bridgman\\Desktop\\Uni\\Projects\\STAT5003\\Data\\Group project\\sum_tab_1.csv"

## TEST DATA ###
file_data_cli_test <- "C:\\Users\\dan bridgman\\Desktop\\Uni\\Projects\\STAT5003\\Data\\Group project\\test_cli.tsv"
file_data_pro_test <- "C:\\Users\\dan bridgman\\Desktop\\Uni\\Projects\\STAT5003\\Data\\Group project\\test_pro.tsv"



# Calling the import_data function from "0 - PrepFunctions.R" for training data
train_cli <- importdata(filetype = "txt",file = file_data_cli,header = TRUE)
train_pro <- importdata(filetype = "txt",file = file_data_pro,header = TRUE)

# Calling the import_data function from "0 - PrepFunctions.R" for test data
test_cli <- importdata(filetype = "txt",file = file_data_cli_test,header = TRUE)
test_pro <- importdata(filetype = "txt",file = file_data_pro_test,header = TRUE)

# Bringing in the labels for training (of which we're not sure which ones are correct and which ones aren't)
train_labels <- importdata(filetype = "csv",file = training_labels,header = TRUE)

```

## Transposing the train & test proteomic files to allow protein strands to become features
## and training numbers to become ids
```{r}

# Transpose the original data frame
t_train_pro <- transpose(train_pro)
colnames(t_train_pro) <- rownames(train_pro)

# Get the sample id to join to other columns
sample <- colnames(train_pro)

# Column bind
t_train_pro <- cbind(sample,t_train_pro)
colnames(t_train_pro)[1] <- "sample"


# Transpose the original data frame
t_test_pro <- transpose(test_pro)
colnames(t_test_pro) <- rownames(test_pro)

# Get the sample id to join to other columns
sample <- colnames(test_pro)

# Column bind
t_test_pro <- cbind(sample,t_test_pro)
colnames(t_test_pro)[1] <- "sample"

t_test_pro
```
# Converting factors to numericals
```{r}
train_cli$gender <- as.numeric(train_cli$gender)
train_cli$msi <- as.numeric(train_cli$msi)

test_cli$gender <- as.numeric(test_cli$gender)
test_cli$msi <- as.numeric(test_cli$msi)

```

# Merging/joining data frames - Cli data, labels and proteomic data
```{r}
# Check ?merge for this section

labels_and_cli <- merge(train_cli,train_labels, by.x = "sample", by.y ="sample", all.x = TRUE )
merged_dfs <- merge(labels_and_cli,t_train_pro, by.x = "sample", by.y ="sample", all.x = TRUE )
#merged_dfs <- merge(merged_dfs,train_labels, by.x = "sample", by.y ="sample", all.x = TRUE )
#df_cleaned <- ImputeData(df, droplist = df[])
merged_dfs
```



```{r}
# Using the Check - Checking missing values and remove, ignore or impute
# How many columns have NAs?
row.has.na <- apply(merged_dfs, 2, function(x){any(is.na(x))})
row.has.na


if (sum(row.has.na) ==0)
  {
    print(" rows have missing values found")
    df_cleaned <- merged_dfs
  } else{
    print(c(sum(row.has.na),"missing values found"))
    print("Histogram of missing merged_dfs")
    aggr_plot <- aggr(merged_dfs, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(merged_dfs), cex.axis=.7, gap=3, ylab=c("Histogram of missing merged_dfs, pre-cleaning","Pattern") )
    missingcols <- colnames(merged_dfs)[colSums(is.na(merged_dfs)) > 0]
    colstoexclude <- colnames(merged_dfs[,-which(names(merged_dfs) %in% missingcols)])
    #Remove NAs
}

```


# Generating some descriptive statistics and some summary stats to check quality of the data and the general shape
```{r}
descriptives <- t(stat.desc(merged_dfs, basic = F))

kable(descriptives, format = "html") %>%
kable_styling(full_width = F)

```

```{r}

```



